---
title: "EP SS prior Debugging"
output: html_notebook
---

```{r}
library(data.table)
library(ggplot2)
```

# Effect of damping

## Small example

```{r small data}
set.seed(1)

v_noise = .1
v_slab = 1
p0 = .2

d = 10
n = 100
X = matrix(rnorm(n * d), nrow = n, ncol=d)
w = c(1, 1, rep(0, d-2))

y = rnorm(n, X %*% w, sqrt(v_noise))

par(mfrow = c(2, 2))
plot(X[, 1], y)
plot(X[, 2], y)
plot(X[, 3], y)
plot(X[, 4], y)
```

## Just a linear model, damping effect

For GSS, converges to same solution regardless of damping:

```{r}
par(mfrow = c(1, 2))

eps_set = c(1, .9, .8, .7, .5, .25)
iters = rep(NA, length(eps_set))
mliks =  rep(NA, length(eps_set))

for (i in 1:length(eps_set)) {
  eps = eps_set[i]
  gss_result = epwpl::GroupSpikeAndSlab(X, y, tau=1/v_noise, p1 = rep(p0, ncol(X)),
                                        v1 = v_slab, verbose=FALSE, opt=FALSE,
                                        damping = eps)
  iters[i] = gss_result$iters
  mliks[i] = gss_result$evidence
}

plot(eps_set, iters)
plot(eps_set, mliks)
```

But my implementation is sensitive to damping:

```{r}
par(mfrow = c(1, 2))

iters = rep(NA, length(eps_set))
mliks =  rep(NA, length(eps_set))

for (i in 1:length(eps_set)) {
  eps = eps_set[i]
  result = epwpl::ep_wlr(X, y, v_noise, v_slab, p0, eps = eps, k=1, opt=FALSE)
  iters[i] = result$iters
  mliks[i] = result$llik
}

plot(eps_set, iters)
plot(eps_set, mliks)
```
My second implementation with factor 2 changes only is insensitive to damping:

```{r}
par(mfrow = c(1, 2))

iters = rep(NA, length(eps_set))
mliks =  rep(NA, length(eps_set))

for (i in 1:length(eps_set)) {
  eps = eps_set[i]
  result = epwpl::ep_ss(X, y, v_noise, v_slab, p0, damping = eps, k=1, opt=FALSE)
  iters[i] = result$iters
  mliks[i] = result$llik
}

plot(eps_set, iters)
plot(eps_set, mliks)
```

## Hyperparam averaging

EP spike and slab linear regression with grid prior over inclusion probability,
Bayesian model averaging.

```{r initial damping}
n_grid = 50
v_noise_grid = rep(v_noise, n_grid)
v_slab_grid = rep(v_slab, n_grid)
p_incl_grid = seq(.01, .9, length.out = n_grid)

eps_set = c(1, .99, .9, .8, .5, .25)
k = 1

par(mfrow = c(2, length(eps_set) / 2))

for (eps in eps_set) {
  is_grid_result = epwpl::ep_grid_ss(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                     opt = FALSE, eps = eps, k = k)
  plot(p_incl_grid, is_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
}


```

We see that the solution is sensitive to initial damping. Meanwhile GSS:

```{r GSS initial damping}
par(mfrow = c(2, length(eps_set) / 2))

gss_result_dt = rbindlist(lapply(eps_set, function(eps) {
  gss_grid_result = epwpl::ep_grid_gss(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                       opt = FALSE, damping = eps, k = k)
  
  plot(p_incl_grid, gss_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
  
  # summarize results for eps setting in dt
  post_incl = data.table(gss_grid_result$alpha)
  setnames(post_incl, format(p_incl_grid, digits = 1))
  post_incl[, dim := 1:nrow(.SD)]
  post_incl = melt(post_incl, id.vars = "dim", variable.name = "p0")
  post_incl[, eps := eps]
  return(post_incl)
}))

ggplot(gss_result_dt, aes(x = dim, y = value, color = as.numeric(p0))) +
  facet_wrap(vars(eps)) +
  geom_point()
```

converges to same solution regardless of initial damping.


With variance optimization:

```{r initial damping with optimization}
n_grid = 50
v_noise_grid = rep(v_noise, n_grid)
v_slab_grid = rep(v_slab, n_grid)
p_incl_grid = seq(.01, .9, length.out = n_grid)

eps_set = c(1, .99, .9, .8, .5, .25)
k = 1

par(mfrow = c(2, length(eps_set) / 2))

for (eps in eps_set) {
  is_grid_result = epwpl::ep_grid_ss(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                     opt = TRUE, eps = eps, k = k)
  plot(p_incl_grid, is_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
}
```

```{r GSS initial damping with optimization}
par(mfrow = c(2, length(eps_set) / 2))

for (eps in eps_set) {
  gss_grid_result = epwpl::ep_grid_gss(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                       opt = TRUE, damping = eps, k = k)
  plot(p_incl_grid, gss_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
}
```

Constant damping:

```{r constant damping}
n_grid = 50
v_noise_grid = rep(v_noise, n_grid)
v_slab_grid = rep(v_slab, n_grid)
p_incl_grid = seq(.01, .9, length.out = n_grid)

eps_set = c(1, .99, .9)
k = 1

par(mfrow = c(2, ceiling(length(eps_set) / 2)))

for (eps in eps_set) {
  is_grid_result = epwpl::ep_grid_lr(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                     opt = TRUE, eps = eps, k = k)
  plot(p_incl_grid, is_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
}
```

## Bigger example

```{r}
d = 100
n = 100
X = cbind(mvtnorm::rmvnorm(n, rep(0, d), 5.*diag(d) + .5*rep(1, d) %*% t(rep(1, d))))
w = c(rep(1, 10), rep(0, d-10))

p0 = .1
v_slab = .1
v_noise = 1/10

y = rnorm(n, X %*% w, sqrt(v_noise))
```

```{r}
n_grid = 50
v_noise_grid = rep(v_noise, n_grid)
v_slab_grid = rep(v_slab, n_grid)
p_incl_grid = seq(.1, .4, length.out = n_grid)

eps_set = c(1, .99, .9)
k = .99

par(mfrow = c(2, ceiling(length(eps_set) / 2)))

for (eps in eps_set) {
  is_grid_result = epwpl::ep_grid_lr(X, y, v_noise_grid, v_slab_grid, qlogis(p_incl_grid),
                                     opt = FALSE, eps = eps, k = k)
  plot(p_incl_grid, is_grid_result$mliks, 
       main = paste0("Initial damping=", eps, ", factor=", k))
}
```


```{r}

par(mfrow = c(2, length(eps_set) / 2))

post_incl = data.table(is_grid_result$alpha)
setnames(post_incl, format(p_incl_grid, digits = 1))
post_incl[, dim := 1:nrow(.SD)]
post_incl = melt(post_incl, id.vars = "dim", variable.name = "p0")

ggplot(post_incl, aes(x = dim, y = value)) +
  geom_point() +
  facet_wrap(vars(p0))

for (eps in eps_set) {
  plot()
}

```



